### 2018-02-22 CO749-GraphColoring/assignments/A3.tex
remains to derive final contradiction

\end{proof}
\end{document}

>>>>

Let $G'$ be the graph obtained from $G$ by identifying vertices $x$ and $y$ (deleting parallel edges if necessary) as vertex $v_{xy}$ and deleting vertex $z$. If $G'$ has a $(k-1)$-coloring $\phi'$ then $\phi'$ can be extending to a $(k-1)$-coloring of $G$ by assigning $\phi'(x ) = \phi'(y) = \phi'(v_{xy})$, then assigning $z$ a color not present on the neighbours of $z$. Such a color exists since $G$ is $(k-1)$-regular and two neighbours of $z$, $x$ and $y$, receive the same color under $\phi'$.
Thus $G'$ is not $(k-1)$-colorable. So $G'$ has a $k$-critical subgraph $K'$. Then $v_{xy} \in V(K')$ for otherwise $K'$ is a $k$-critical strict subgraph of $G$, contradicting that $G$ is $k$-critical. By minimality of $G$ we have
$$p(K') \leq 0$$
with equality if and only if $K'$ is $K_k$. Let $K$ be the graph $K' - v_{xy} + \{x,y\}$. Then $K$ is a strict subgraph of $G$, since $z \not\in V(K)$. So by the previous claim
$$k-1 \leq p(G[V(K')]) \leq p(K')$$
Now we observe that
$$p(K) = p(K') + (k-1) - 2|E(v_{xy}, N(x) \cap N(y) \cap V(K'))|$$
Since $K$ has more vertex that $K'$, and parallel edges were deleted in obtaining $G'$. Combining inequalities we see that
$$k-1 \leq p(K) \leq 0 + (k-1) - 2|E(v_{xy}, N(x) \cap N(y) \cap V(K'))|$$
and hence 
$$|E(v_{xy}, N(x) \cap N(y) \cap V(K'))| \leq 0.$$
Since $|E(v_{xy}, N(x) \cap N(y) \cap V(K'))|$ is non-negative this implies $|E(v_{xy}, N(x) \cap N(y) \cap V(K'))| = 0$ and thus
$N(x)\cap N(y) \cap V(K') = \emptyset$. Furthermore the displayed inequalities are tight, implying that $K'$ is isomorphic to $K_k$. Letting $H = K'-v_{xy}$ we obtain the desired clique of the claim.

\end{document}
